SPP-Perc: Peceptron Learning Based Enhanced Signature Path Prefetcher

### 2018-11-16 method.tex
Incorporating suggestions

\section{METHODOLOGY}
\subsection{Performance Model}
For testing and comparing SPP-Perc, we used ChampSim\cite{Champsim}
simulator.  ChampSim is an enhanced version of the framework that was
used for the 2nd Data Prefetch Championship (DPC-2).\cite{DPC_2} We
model model a 1-core / 4-core / 8-core Out of Order machine. The
details of the configuration parameters is summarized in Table
\ref{tab:Sim_params}.

@@

\subsection{Testing Under Additional Memory Constraints}
The default single-core configuration simulates a 2MB LLC and a single
channel DRAM with 12.8GB/s bandwidth.  We extend the simulations to
include memory constraints introduced in DPC-2.  Specifically we look
at the following two variations:
\begin{itemize}
\item \textit{Low Bandwidth DRAM}: Here the DRAM bandwidth is limited
  to 3.2 GB/s
\item \textit{Small LLC}: In this scenario, LLC size is reduced to 512
  KB
\end{itemize}
All the multi-core simulations are only done in the default
configuration.

\subsection{Workloads}
This is the first time that SPEC 2017 benchmark suite\cite{SPEC2017}
has been used to characterize and measure the prefetch performance.
We use all the 20 workloads available in the SPEC 2017 Suite.  Using
SimPoint\cite{SimPoint} methodology, we identified 95 different
program segments of 1 Billion instructions each.

We performed cross-validation of our SPP-Perc model using SPEC
2006\cite{SPEC2006} and CloudSuite\cite{CloudSuite} benchmarks.  For
SPEC 2006, we developed 94 simpoints spread across the 29
applications.  For CloudSuite, we used the traces made available for
the 2nd Cache Replacement Competition (CRC-2)\cite{CRC_2}.  The traces
include 4 multi-core applications with 6 distinct phases per
application.  In total, we used 285 traces representing workloads
across 53 applications.

@@

\subsection{Preferchers Simulated}
We compared SPP-Perc against three of the latest state of the art
prefetchers: Best Offset Prefetcher (BOP), DRAM Aware - Access Map
Pattern Matching (DA-AMPM)\cite{DA_AMPM} and Signature Path Prefecher
(SPP).  BOP was the winner of 2nd Data Prefetching Championship.
DA-AMPM is the enhanced version of AMPM, modified to account for DRAM
row buffer locality.  SPP has been shown to outperform BOP on SPEC
2006 traces.  For each of these, we compare their speedups taking no
prefetching as the baseline.

\subsection{Developing Features for SPP-Perc}
This section describes the intuition and analysis that went behind
finalizing the features.  As noted earlier, we developed a set of 9
features which allow the perceptron throttler to correlate prefetching
decision with the program context.  To study the correlation across
each feature, we study statistically the perceptron weights and try to
interpret their distribution.

@@

As a part of feature selection, we initially came up a mix of 23
features- primary and composite.  By studying cross correlation of
each of these features against others in a 23x23 matrix, we identified
pairs of features with correlation factor > 0.9 in magnitude and
carefully eliminated redundant features.  Using this approach, we
managed to reduce the feature count to 9.  Thus, in the final
implementation of SPP-Perc, no two features have a high correlation
between them.  This way we can be sure that each feature is bringing
in contribution that cannot be captured using other features.

@@

\subsection{Overhead for SPP-Perc}

@@

\begin{table}[]
    \centering
    \begin{tabular}{|p{1.5cm}|p{1cm}|c|p{1.5cm}|}
    \hline
        \textbf{Structure} &
        \textbf{Entry} &
        \textbf{Components} &
        \textbf{Total} \\
    \hline
        \multirow{5}{1.5cm}{Signature Table} & \multirow{5}{1.5cm}{256} & Valid (1 bit) & \multirow{5}{1cm}{11008  bits}\\
        & & Tag (16 bits) &\\
        & & Last Offset (6 bits) &\\  
        & & Signature (12 bits) &\\
        & & LRU (6 bits) &\\
    \hline
        \multirow{3}{1.5cm}{Pattern Table} & \multirow{3}{1.5cm}{512} & $C_{sig}$ (4bits) &\multirow{3}{1cm}{24576 bits}\\
        & & $C_{delta}$ (4*4 bits) &\\
        & & Delta (4*7 bits) &\\
    \hline
        \multirow{3}{1.5cm}{Perceptron Tables} & \multirow{3}{1.5cm}{4096*8 128*1} & \multirow{3}{1.5cm}{5 bits} & \multirow{3}{1.5cm}{164480 bits}\\
        & & &\\
        & & &\\
    \hline
        \multirow{2}{1.5cm}{Prefetch Filter \footnotemark[1]} & \multirow{2}{1.5cm}{1024} & \multirow{2}{1.5cm}{84 bits} & \multirow{2}{1.5cm}{86016 bits}\\
        & & &\\
    \hline
     \multirow{2}{1.5cm}{Reject Filter \footnotemark[2]} & \multirow{2}{1.5cm}{1024} & \multirow{2}{1.5cm}{83 bits} & \multirow{2}{1.5cm}{84992 bits}\\
        & & &\\
    \hline
        \multirow{4}{1.5cm}{Global History Register} & \multirow{4}{1.5cm}{8} & Signature (12 bits) & \multirow{4}{1.5cm}{264 bits}\\
        & & Confidence (8 bits) & \\
        & & Last Offset (6 bits) & \\
        & & Delta (7 bits) & \\
    \hline
        \multirow{2}{1.5cm}{Accuracy Counters} & 1 & C$_{total}$ & 10 bits\\
        & 1 & C$_{useful}$ & 10 bits\\
    \hline
        \multirow{3}{1.5cm}{Global PC Trackers} & \multirow{3}{1cm}{3} & $PC_1$ (12 bits)& \multirow{3}{1.cm}{36 bits}\\
        & & $PC_2$ (12 bits) &\\
        & & $PC_3$ (12 bits) &\\
    \hline
        \multicolumn{4}{|c|}{\textbf{Total: 371356 bits = 45.33 KB}}\\
    \hline
    \end{tabular}
    \caption{SPP-Perc Storage Overhead}
    \label{tab:SPPPerc_overhead}
\end{table}
\footnotetext[1]{Components of Prefetch Filter can be found in Table \ref{tab:PF_metadata}}
\footnotetext[2]{RF does not need to maintain the useful bit as that only applies for prefetches that ultimately made through}

In this section, we analyze the hardware overhead required to
implement SPP-Perc.  The Prefetch Filter was enhanced to accommodate
storing of meta-data for perceptron training.  Table
\ref{tab:PF_metadata} depicts the meta-data stored for each entry in
the Prefetch Filter.  Table \ref{tab:SPPPerc_overhead} shows the total
storage overhead of SPP-Perc implementation.  The hardware budget for
2nd Data Prefetching championship was 32 KB.  Keeping that in mind and
the speedup SPP-Perc obtained over the winner, the extra hardware
budget can be accounted for.  The extra hardware also makes the
overall scheme more scalable than SPP.  In the original SPP paper, it
was demonstrated that adding extra hardware brings little advantage in
terms of performance gain.  The newly added perceptron tables can be
scaled to increase / decrease features depending on the permitted
budget.

>>>>

\section{METHODOLOGY}
\label{Method}

\subsection{Performance Model}
\label{Method-Model}
For testing and comparing NATCH, we used ChampSim\cite{Champsim}
simulator.  ChampSim is an enhanced version of the framework that was
used for the 2nd Data Prefetch Championship (DPC-2).\cite{DPC_2} We
model model a 1-core / 4-core / 8-core Out of Order machine. The
details of the configuration parameters is summarized in Table
\ref{tab:Sim_params}.

@@

\subsection{Testing Under Additional Memory Constraints}
\label{Method-AdditionalMem}
The default single-core configuration simulates a 2MB LLC and a single
channel DRAM with 12.8GB/s bandwidth.  We extend the simulations to
include memory constraints introduced in DPC-2.  Specifically we look
at the following two variations:
\begin{itemize}
\item \textit{Low Bandwidth DRAM}: Here the DRAM bandwidth is limited
  to 3.2 GB/s
\item \textit{Small LLC}: In this scenario, LLC size is reduced to 512
  KB
\end{itemize}
All the multi-core simulations are only done in the default
configuration.

\subsection{Workloads}
\label{Method-Workloads}
This is the first time that SPEC 2017 benchmark suite\cite{SPEC2017}
has been used to characterize and measure the prefetch performance.
We use all the 20 workloads available in the SPEC 2017 Suite.  Using
SimPoint\cite{SimPoint} methodology, we identified 95 different
program segments of 1 Billion instructions each.

We performed cross-validation of our NATCH model using SPEC
2006\cite{SPEC2006} and CloudSuite\cite{CloudSuite} benchmarks.  For
SPEC 2006, we developed 94 simpoints spread across the 29
applications.  For CloudSuite, we used the traces made available for
the 2nd Cache Replacement Competition (CRC-2)\cite{CRC_2}.  The traces
include 4 multi-core applications with 6 distinct phases per
application.  In total, we used 285 traces representing workloads
across 53 applications.

@@

\subsection{Preferchers Simulated}
\label{Method-Prefetchers}
We compared NATCH against three of the latest state of the art
hardware-only prefetchers: Best Offset Prefetcher (BOP), DRAM Aware - Access Map
Pattern Matching (DA-AMPM)\cite{DA_AMPM} and Signature Path Prefecher
(SPP).  BOP was the winner of 2nd Data Prefetching Championship.
DA-AMPM is the enhanced version of AMPM, modified to account for DRAM
row buffer locality.  SPP has been shown to outperform BOP on SPEC
2006 traces.  For each of these, we compare their speedups taking no
prefetching as the baseline.

\subsection{Developing Features for NATCH}
\label{Method-Features}
This section describes the intuition and analysis that went behind
finalizing the features.  As noted earlier, we developed a set of 9
features which allow the perceptron throttler to correlate prefetching
decision with the program context.  To study the correlation across
each feature, we study statistically the perceptron weights and try to
interpret their distribution.

@@

As a part of feature selection, we initially came up a mix of 23
features- primary and composite.  By studying cross correlation of
each of these features against others in a 23x23 matrix, we identified
pairs of features with correlation factor > 0.9 in magnitude and
carefully eliminated redundant features.  Using this approach, we
managed to reduce the feature count to 9.  Thus, in the final
implementation of NATCH, no two features have a high correlation
between them.  This way we can be sure that each feature is bringing
in contribution that cannot be captured using other features.

@@

\subsection{Overhead for NATCH}
\label{Method-Overheads}

@@

\begin{table}[]
    \centering
    \begin{tabular}{|p{1.5cm}|p{1cm}|c|p{1.5cm}|}
    \hline
        \textbf{Structure} &
        \textbf{Entry} &
        \textbf{Components} &
        \textbf{Total} \\
    \hline
        \multirow{5}{1.5cm}{Signature Table} & \multirow{5}{1.5cm}{256} & Valid (1 bit) & \multirow{5}{1cm}{11008  bits}\\
        & & Tag (16 bits) &\\
        & & Last Offset (6 bits) &\\  
        & & Signature (12 bits) &\\
        & & LRU (6 bits) &\\
    \hline
        \multirow{3}{1.5cm}{Pattern Table} & \multirow{3}{1.5cm}{512} & $C_{sig}$ (4bits) &\multirow{3}{1cm}{24576 bits}\\
        & & $C_{delta}$ (4*4 bits) &\\
        & & Delta (4*7 bits) &\\
    \hline
        \multirow{3}{1.5cm}{Perceptron Tables} & \multirow{3}{1.5cm}{4096*8 128*1} & \multirow{3}{1.5cm}{5 bits} & \multirow{3}{1.5cm}{164480 bits}\\
        & & &\\
        & & &\\
    \hline
        \multirow{2}{1.5cm}{Prefetch Filter \footnotemark[1]} & \multirow{2}{1.5cm}{1024} & \multirow{2}{1.5cm}{84 bits} & \multirow{2}{1.5cm}{86016 bits}\\
        & & &\\
    \hline
     \multirow{2}{1.5cm}{Reject Filter \footnotemark[2]} & \multirow{2}{1.5cm}{1024} & \multirow{2}{1.5cm}{83 bits} & \multirow{2}{1.5cm}{84992 bits}\\
        & & &\\
    \hline
        \multirow{4}{1.5cm}{Global History Register} & \multirow{4}{1.5cm}{8} & Signature (12 bits) & \multirow{4}{1.5cm}{264 bits}\\
        & & Confidence (8 bits) & \\
        & & Last Offset (6 bits) & \\
        & & Delta (7 bits) & \\
    \hline
        \multirow{2}{1.5cm}{Accuracy Counters} & 1 & C$_{total}$ & 10 bits\\
        & 1 & C$_{useful}$ & 10 bits\\
    \hline
        \multirow{3}{1.5cm}{Global PC Trackers} & \multirow{3}{1cm}{3} & $PC_1$ (12 bits)& \multirow{3}{1.cm}{36 bits}\\
        & & $PC_2$ (12 bits) &\\
        & & $PC_3$ (12 bits) &\\
    \hline
        \multicolumn{4}{|c|}{\textbf{Total: 371356 bits = 45.33 KB}}\\
    \hline
    \end{tabular}
    \caption{NATCH Storage Overhead}
    \label{tab:NATCH_overhead}
\end{table}
\footnotetext[1]{Components of Prefetch Filter can be found in Table \ref{tab:PF_metadata}}
\footnotetext[2]{RF does not need to maintain the useful bit as that only applies for prefetches that ultimately made through}

In this section, we analyze the hardware overhead required to
implement NATCH.  The Prefetch Filter was enhanced to accommodate
storing of meta-data for perceptron training.  Table
\ref{tab:PF_metadata} depicts the meta-data stored for each entry in
the Prefetch Filter.  Table \ref{tab:NATCH_overhead} shows the total
storage overhead of NATCH implementation.  The hardware budget for
2nd Data Prefetching championship was 32 KB.  Keeping that in mind 
the considerable speedup NATCH obtained over the winner, the extra hardware
budget can be accounted for.  The extra hardware also makes the
overall scheme more scalable than SPP.  In the original SPP paper, it
was demonstrated that adding extra hardware brings little advantage in
terms of performance gain.  The newly added perceptron tables can be
scaled to increase / decrease features depending on the permitted
budget.
